A Survey of Quantitative Descriptions of Molecular Structure

### 2012-04-01 paper.tex
Added more text ondescriptor types

A number of geometric descriptors characterize the molecular
surface. The simplest such example is the surface area descriptor (and
the corresponding volume descriptor). A simple extension of such a
descriptor is to characterize distributions of a physicochemical
property over the molecular surface. Note that surface descriptor
calculations based onanalytical \cite{Connolly:1983aa} or tesselation
algorithms \cite{Eisenhaber:1995qf} can be slow and that parametrized
methods such as the VSA descriptors \cite{Labute:2008aa} that are
based on precalculated surface area values derived from a list of
functional groups. A comparison of surface area values generated by
the two approaches for XXX molecules indicates a high degree of
correlation.

@@

Fingerprints are a compact representation of a chemical structure and
are used in a variety of ways ranging from database searches looking
for similar structures or substructures 
to virtual screenings studies
\cite{Sun:2006kx}. In many search applications, the binary nature of
fingerprints allow for efficient algorithms to compare fingerprints
and a number of heuristics have been developed that allow for rapid
similarity searches \cite{Baldi:2008aa,Swamidass:2007ab}.

@@

\section{Running Notes}
\begin{itemize}
\item \textbf{Done} What are descriptors, why are they needed?
\item Survey of descriptor types, focusing on nature of inputs,
  complexity, interpretabilty 
\item \textbf{Done} Talk about biological descriptors (see paper from Tropsha that
  uses bioactivity data as descriptors)
\item Discuss how fingerprints can also be used as descriptors
\item Also mention descriptors for non-small molecule cases
  \begin{itemize}
  \item proteins
  \item polymers
  \item inorganics (maybe ask Maciej for pointers)
  \end{itemize}
\item Survey of descriptor implementations (OSS and non-OSS)
  \begin{itemize}
  \item MOE
  \item DRAGON
  \item CDK
  \item RDKit
  \item JChem
  \item PipelinePilot
  \item OEChem/GraphTK
  \item JOELib
  \item CODESSA
  \item Schrodinger stuff (Meastro?)
  \end{itemize}
\item Go into more detail of OSS descriptor implementations
  \begin{itemize}
  \item CDK (describe architecture), mention wrappers around the
    library (CDKDescUI and batch mode)
  \item PaDEL
  \end{itemize}
\item Compare OSS descs to commercial ones - point to the Pfizer/CDD
  paper that compared MOE and CDK
\end{itemize}
\clearpage
\newpage

>>>>

A number of geometric descriptors characterize the molecular
surface. The simplest such example is the surface area descriptor (and
the corresponding volume descriptor). A simple extension of such a
descriptor is to characterize distributions of a physicochemical
property over the molecular surface such as the Charged Partial
Surface Area descriptors \cite{Stanton:1990aa} that characterize the
partial charge distribution over the molecular surfce. Note that
surface descriptor calculations based on analytical
\cite{Connolly:1983aa} or tesselation algorithms
\cite{Eisenhaber:1995qf} can be slow and that parametrized methods
such as the VSA descriptors \cite{Labute:2008aa} that are based on
precalculated surface area values derived from a list of functional
groups. A comparison of surface area values generated by the two
approaches for XXX molecules indicates a high degree of correlation.
\textbf{Ray tracing descriptors, ultrafast shape descriptors,
  spherical harmonic shape descriptors, field based descriptors
  [Cresset etc]}

@@

Fingerprints are a compact representation of a chemical structure and
are used in a variety of ways ranging from database searches looking
for similar structures or substructures to virtual screenings studies
\cite{Sun:2006kx}. In many search applications, the binary nature of
fingerprints allow for efficient algorithms to compare molecules and
evaluate similarity and a number of heuristics have been developed
that allow for rapid similarity searches
\cite{Baldi:2008aa,Swamidass:2007ab}. In virtual screening scenarios,
fingerprints can be used as features. In many cases, the entire
fingerprint is employed as a set of independent variables. Since this
can lead to poor predictive models, some sort of feature selection is
usually employed and there have been studies that address the problem
of identifying subsets of bits from a fingerprint that lead to optimal
performance in predictive models (albeit, usually in a target-class
specific manner) \cite{Xue:2004fk}.

@@

\section{Running Notes}
\begin{itemize}
\item \textbf{Done} What are descriptors, why are they needed?
\item Survey of descriptor types, focusing on nature of inputs,
  complexity, interpretabilty 
\item \textbf{Done} Talk about biological descriptors (see paper from Tropsha that
  uses bioactivity data as descriptors)
\item \textbf{Done} Discuss how fingerprints can also be used as descriptors
\item Also mention descriptors for non-small molecule cases
  \begin{itemize}
  \item proteins
  \item polymers \& mixtures
  \item \textbf{Done} inorganics (maybe ask Maciej for pointers)
  \end{itemize}
\item Survey of descriptor implementations (OSS and non-OSS)
  \begin{itemize}
  \item MOE
  \item DRAGON
  \item CDK
  \item RDKit
  \item JChem
  \item PipelinePilot
  \item OEChem/GraphTK
  \item JOELib
  \item CODESSA
  \item Schrodinger stuff (Meastro?)
  \end{itemize}
\item Go into more detail of OSS descriptor implementations
  \begin{itemize}
  \item CDK (describe architecture), mention wrappers around the
    library (CDKDescUI and batch mode)
  \item PaDEL
  \end{itemize}
\item Compare OSS descs to commercial ones - point to the Pfizer/CDD
  paper that compared MOE and CDK
\end{itemize}
\clearpage
\newpage
